---
title: "Lit_Review_Analysis"
author: "Sarah Roberts"
date: "7/1/2022"
output: html_document
---

#1. Prep
Load important libraries 
```{r}
library(tidyverse)
library(ggplot2)
```




load data 
```{r}
dat <- readxl::read_xlsx("Data/DataExtraction_Jacoby.xlsx")


```

clean it up a bit
```{r} 
dat <- dat %>% 
  dplyr::select(-"...1") %>% #remove first column
  rename(Unique_ID = "...2") %>% #rename column
  filter(!row_number() %in% 1) #remove first row

```

##decision point 1
remove rows where data type was none
```{r}
dat <- dat %>% filter(`Data Type` != c("none", "isotope")) #isotope just tells us what they are similar to 
```



clean up prey column to be long format 
```{r}
dat$Prey <- gsub(pattern = ',', replacement = ";", x = dat$Prey) #make all commmas semi colon 
dat <- dat %>% separate_rows(Prey, sep=';')
```

clean up weird spacing at beg and end of Prey column
```{r}
dat <- dat %>%
  mutate_if(is.character, str_trim) #removes whitespace at beginning and end of all characters
```

which prey are named different things?
```{r}
tallied <- dat %>% group_by(Prey) %>% tally()
tallied$Prey
```
okay - we need to combine things that are named differently into one prey item. 

```{r}
dat$Prey <- ifelse(dat$Prey == 'anchovies'| dat$Prey == "Fatty acid: anchovy", "anchovy spp (Anchoa spp.)", 
                   ifelse(dat$Prey == 'and crabs', "crabs", 
                          ifelse( dat$Prey == 'and oystr toadfish'| dat$Prey == "toadfish (Opsanus tau)", "Oyster toadfish (Opsanus tau)", 
                                  ifelse(dat$Prey == "Atlantic croaker"| dat$Prey == "croacker"| dat$Prey == "croaker"| dat$Prey == "Croaker Sciaenid", "Atlantic croaker (Micropogonias undulatus)",
                                         ifelse(dat$Prey == "Bluefish Pomatomus saltatrix", "Bluefish (Pomatomus saltatrix)", 
                                                ifelse(dat$Prey == "brief squid"| dat$Prey == "Brief squid Lolliguncula brevis", "Atlantic brief squid (Lolliguncula brevis)", 
                                                       ifelse(dat$Prey == "Congridae Conger", "Conger eel (Conger oceanicus)", 
                                                              ifelse(dat$Prey == "Decapoda \"Shrimp\""| dat$Prey == "Shrimp Pernaeus sp.", "shrimp", 
                                                                     ifelse(dat$Prey == "Enoploteuthis sp.", "Enoploteuthis",
                                                                            ifelse(dat$Prey == "grunt"| dat$Prey == "Grunt Haemulid"| dat$Prey == "Haemulon", "Grunt Haemulon sp.", 
                                                                                   ifelse(dat$Prey == "hakee spp. (Urophycis spp.)", "Hake (Urophycis sp.)", ifelse(dat$Prey == "Histioteuthis", "Histioteuthis cr H. miranda", ifelse( dat$Prey == "Illex illecebrosus", "Illex spp.", ifelse(dat$Prey == "Labrid", "Labridae", ifelse(dat$Prey == "ladyfish", "ladyfish (Elops saurus)", dat$Prey)))))))))))))))

 dat <- dat %>% dplyr::filter(Prey %!in% c("fish", "Fish", "Fish (1)"))
```

probably should create a column called species order
#plotting

```{r}
dat %>% 
  dplyr::filter(Ecotype == "Coastal")%>% 
  ggplot(aes(x=Prey)) + 
  geom_bar(aes(y = ..count..)) + theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
```
