---
title: "Lit_Review_Analysis"
author: "Sarah Roberts"
date: "7/1/2022"
output: html_document
---

#1. Prep
Load important libraries 
```{r}
library(tidyverse)
library(ggplot2)
library(readxl)
library(ggthemes)
library(scales)
library(gridExtra)

theme_Publication <- function(base_size=12, base_family="Arial") {

      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size= unit(0.5, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0")
          ))
      
}

Mode <- function(x) {
 ux <- unique(x)
 ux[which.max(tabulate(match(x, ux)))]
}
```




load data 
```{r}
dat <- readxl::read_xlsx("Data/DataExtraction_Jacoby_20220727_TestforSarah.xlsx")


```

clean it up a bit
```{r} 
dat <- dat %>% 
  dplyr::select(-"...1") %>% #remove first column
  filter(!row_number() %in% 1) #remove first row

```

##decision point 1
remove rows where data type was none
```{r}
dat <- dat %>% filter(Data_Type != c("none")) #isotope just tells us what they are similar to - should we remove?  
```



clean up prey column to be long format 
```{r}

dat <- dat %>% separate_rows(New_Prey, sep=' ')
```

clean up weird spacing at beg and end of Prey column
```{r}
dat <- dat %>%
  mutate_if(is.character, str_trim) #removes whitespace at beginning and end of all characters
```

#add in prey order 
```{r}
prey_info <- readxl::read_xlsx("Data/PreyList.xlsx")

dat <- left_join(dat, prey_info, by = c("New_Prey" = "Prey"))
```

#plotting

by family and region 
```{r}
#try a trick to reorder each individual facet
plot_dat <- dat %>% group_by(Ecotype, Region, Prey_Species, Prey_Family) %>% tally()

plot_dat %>% 
  dplyr::filter(Ecotype != "Unknown") %>%
  ggplot(aes(reorder_within(Prey_Family, n, Region), n)) +
   geom_bar(stat = 'identity') +
  scale_x_reordered() +
  facet_wrap(~ Region, scales = "free_y")+
    coord_flip()

plot_dat %>% 
  dplyr::filter(Ecotype != "Unknown" & Prey_Family != "NA") %>%
  ggplot(aes(reorder_within(Prey_Family, n, list(Ecotype, Region)), n)) +
   geom_bar(stat = 'identity') +
  scale_x_reordered() +
  facet_wrap(~ Ecotype + Region, scales = "free")+ theme_Publication()+
    coord_flip() + ggtitle("Dominant Prey Family by Ecotype and Region")


#try it as a proportion 
plot_dat <- dat %>% 
  group_by(Ecotype, Region, Prey_Family) %>% 
  summarise(n = n()) %>% 
  mutate(freq = (n/sum(n))*100)
  
p <- plot_dat %>% 
  dplyr::filter(Ecotype != "Unknown") %>%
  ggplot(aes(reorder_within(Prey_Family, freq, list(Ecotype, Region)), freq)) +
   geom_bar(stat = 'identity', fill = "orange") +
  scale_x_reordered() +
  facet_wrap(~ Ecotype + Region, scales = "free_y")+
    labs(y = "Percentage of studies", x = "Prey Family", lab.nb.digits=1, orientation = "horiz", lab.size = 7, lab.hjust=-1, xlim=c(0,100))+ theme_Publication()+
    coord_flip() + ggtitle("Dominant Prey Family by Ecotype and Region")
p
ggsave(filename= "Figures/Lit_review_Family_Region_all.jpg", plot=p, width = 40, height=20, units=c("cm"), dpi=500)


#just get out the top 10 or so 
p <- plot_dat %>% top_n(10, wt = freq) %>% 
  dplyr::filter(Ecotype != "Unknown") %>%
  ggplot(aes(reorder_within(Prey_Family, freq, list(Ecotype, Region)), freq)) +
   geom_bar(stat = 'identity', fill = "orange") +
  scale_x_reordered() +
  facet_wrap(~ Ecotype + Region, scales = "free_y")+
    labs(y = "Percentage of studies", x = "Prey Family", lab.nb.digits=1, orientation = "horiz", lab.size = 7, lab.hjust=-1, xlim=c(0,100))+ theme_Publication()+
    coord_flip() + ggtitle("Top 10 Prey Family by Ecotype and Region")
p
ggsave(filename= "Figures/Lit_review_Family_Region_10.jpg", plot=p, width = 40, height=20, units=c("cm"), dpi=500)

```

by family and location (state)

```{r}
plot_dat <- dat %>% 
  group_by(Ecotype, Location, Prey_Family) %>% 
  summarise(n = n()) %>% 
  mutate(freq = (n/sum(n))*100)
  
p <- plot_dat %>% 
  dplyr::filter(Ecotype != "Unknown") %>%
  ggplot(aes(reorder_within(Prey_Family, freq, list(Ecotype, Location)), freq)) +
   geom_bar(stat = 'identity', fill = "orange") +
  scale_x_reordered() +
  facet_wrap(~ Ecotype + Location, scales = "free_y")+
    labs(y = "Percentage of studies", x = "Prey Family", lab.nb.digits=1, orientation = "horiz", lab.size = 7, lab.hjust=-1, xlim=c(0,100))+ theme_Publication()+
    coord_flip() + ggtitle("Dominant Prey Family by Ecotype and Location")
p
ggsave(filename= "Figures/Lit_review_Family_Location_all.jpg", plot=p, width = 40, height=20, units=c("cm"), dpi=500)

#just get out the top 10 or so 
p <- plot_dat %>% top_n(10, wt = freq) %>% 
  dplyr::filter(Ecotype != "Unknown") %>%
  ggplot(aes(reorder_within(Prey_Family, freq, list(Ecotype, Location)), freq)) +
   geom_bar(stat = 'identity', fill = "orange") +
  scale_x_reordered() +
  facet_wrap(~ Ecotype + Location, scales = "free_y")+
    labs(y = "Percentage of studies", x = "Prey Family", lab.nb.digits=1, orientation = "horiz", lab.size = 7, lab.hjust=-1, xlim=c(0,100))+ theme_Publication()+
    coord_flip() + ggtitle("Top 10 Prey Family by Ecotype and Location")
p
ggsave(filename= "Figures/Lit_review_Family_Location_10.jpg", plot=p, width = 40, height=20, units=c("cm"), dpi=500)

```

by species and region 
```{r}
#do top species by region 
plot_dat <- dat %>% 
  group_by(Ecotype, Region, Prey_Species) %>% 
  summarise(n = n()) %>% 
  mutate(freq = (n/sum(n))*100)

p <- plot_dat %>% 
  dplyr::filter(Ecotype != "Unknown") %>%
  ggplot(aes(reorder_within(Prey_Species, freq, list(Ecotype, Region)), freq)) +
   geom_bar(stat = 'identity', fill = "orange") +
  scale_x_reordered() +
  facet_wrap(~ Ecotype + Region, scales = "free_y")+
    labs(y = "Percentage of studies", x = "Prey Species", lab.nb.digits=1, orientation = "horiz", lab.size = 7, lab.hjust=-1, xlim=c(0,100))+ theme_Publication()+
    coord_flip() + ggtitle("Dominant Prey Species by Ecotype and Region")
p
ggsave(filename= "Figures/Lit_review_Species_Region_all.jpg", plot=p, width = 40, height=20, units=c("cm"), dpi=500)

#just get out the top 10 or so 
p <- plot_dat %>% top_n(10, wt = freq) %>% 
  dplyr::filter(Ecotype != "Unknown") %>%
  ggplot(aes(reorder_within(Prey_Species, freq, list(Ecotype, Region)), freq)) +
   geom_bar(stat = 'identity', fill = "orange") +
  scale_x_reordered() +
  facet_wrap(~ Ecotype + Region, scales = "free_y")+
    labs(y = "Percentage of studies", x = "Prey Species", lab.nb.digits=1, orientation = "horiz", lab.size = 7, lab.hjust=-1, xlim=c(0,100))+ theme_Publication()+
    coord_flip() + ggtitle("Top 10 Prey Species by Ecotype and Region")
p
ggsave(filename= "Figures/Lit_review_Species_Region_10.jpg", plot=p, width = 40, height=20, units=c("cm"), dpi=500)
```

#summary stats 
how many studies were there, how many for each ecotype, region, how many identified to what level 

Okay - we need to talk about what we want to summarise here because some studies are associated with multiple regions or multiple 
```{r}
dat_orig <- readxl::read_xlsx("Data/DataExtraction_Jacoby_20220727_TestforSarah.xlsx")
dat_orig <- dat_orig %>% 
  dplyr::select(-"...1") %>% #remove first column
  filter(!row_number() %in% 1) #remove first row

dat_orig <- dat_orig %>% filter(Data_Type != c("none"))
dat_orig <- dat_orig %>% dplyr::select(-Old_Prey, -New_Prey)

#so we could read this as there were at least 4 studies that focused on the east coast (while some studies could focus on both the east coast and mid atlantic)
summary(as.factor(dat_orig$Region))
summary(as.factor(dat_orig$Ecotype))
summary(as.factor(dat_orig$Data_Type))

#nope, because what if it is counting one study twice 

#okay this might just do it 
Data_Type <- dat_orig %>% group_by(Data_Type) %>% summarise(Number_of_Studies = n_distinct(ID))
Region <- dat_orig %>% group_by(Region) %>% summarise(Number_of_Studies = n_distinct(ID))
Ecotype <- dat_orig %>% group_by(Ecotype) %>% summarise(Number_of_Studies = n_distinct(ID))
Level <- dat %>% group_by(Prey_Classification)%>% summarise(Number_of_Studies = n_distinct(ID))
Level[is.na(Level)] <- "None"
Source <- dat_orig %>% group_by(Source) %>% summarise(Number_of_Studies = n_distinct(ID))

#something can be in two groups - which is why the sums dont add up! 
n_distinct(dat_orig$ID)
sum(Data_Type$Number_of_Studies)
sum(Region$Number_of_Studies)
sum(Ecotype$Number_of_Studies)
sum(Source$Number_of_Studies)
sum(Level$Number_of_Studies)

#plot those things out 

p1 <- Data_Type %>% 
  ggplot(aes(reorder(Data_Type, Number_of_Studies), Number_of_Studies)) +
   geom_bar(stat = 'identity') +
  scale_x_reordered()+ theme_Publication()+
    labs(y = "Number of Studies", x = "Data Type")+
    coord_flip()

p2 <- Region %>% 
  ggplot(aes(reorder(Region, Number_of_Studies), Number_of_Studies)) +
   geom_bar(stat = 'identity') +
  scale_x_reordered()+ theme_Publication()+
    labs(y = "Number of Studies", x = "Region")+
    coord_flip()

p3 <- Ecotype %>% 
  ggplot(aes(reorder(Ecotype, Number_of_Studies), Number_of_Studies)) +
   geom_bar(stat = 'identity') +
  scale_x_reordered()+ theme_Publication()+
    labs(y = "Number of Studies", x = "Ecotype")+
    coord_flip()

p4 <- Level %>% 
  ggplot(aes(reorder(Prey_Classification, Number_of_Studies), Number_of_Studies)) +
   geom_bar(stat = 'identity') +
  scale_x_reordered()+ theme_Publication()+
    labs(y = "Number of Studies", x = "Classification Level")+
    coord_flip()

p5 <- Source %>% 
  ggplot(aes(reorder(Source, Number_of_Studies), Number_of_Studies)) +
   geom_bar(stat = 'identity') +
  scale_x_reordered()+ theme_Publication()+
    labs(y = "Number of Studies", x = "Source")+
    coord_flip()

a <- grid.arrange(p1, p2, p3, p4, p5, nrow = 2)
ggsave(filename= "Figures/Lit_review_stats.jpg", plot=a, width = 30, height=20, units=c("cm"), dpi=500)
```